Predicting Transport in Lyotropic Liquid Crystal Membranes with Molecular Dynamics Simulations

### 2017-08-18 Ben_Manuscripts/Masters_Bypass/methods.tex
polished

Initial guesses for the remaining structural parameters were chosen
based on experimental data and treated as variables during model
development. The distance between pores was based on experimental SAXS
data for this system \cite{feng_thin_2016}. Initial configuration pore
spacing was chosen to be 10 \% larger than the experimental value to reduce
unintended repulsions resulting from a tightly packed initial configuration. 
The layer spacing was based on experimental 2D WAXS data which shows
reflections corresponding to features spaced 3.7 \angstrom apart. It has 
been hypothesized that the features are present due to pi-pi
interactions between stacked aromatic rings \cite{feng_scalable_2014}. 
Our simulations tend to equilibrate to a wider interlayer spacing of $\approx$ 4.1 \angstrom,
which inspired separate systems starting with layer spacings greater than
 4 \angstrom. We estimate the pore radius to be 0.6 nm based on past TEM 
images and size exclusion rejection data\cite{feng_scalable_2014,
feng_thin_2016,zhou_supported_2005}. Comparing a geometric
measurement of pore size taken from an atomistic model, to a less
precise, experimentally derived pore size estimate, will give ambiguous
results. When constructing pores, we chose the carboxylate carbon from
the monomer head group as a reference atom, and placed it a distance r from the
pore center, where r is the pore radius. We will not make direct comparisons
of pore radius between our model and experiment to avoid the ambiguity. 
% There is plenty I can say about initial configuration. Starting layers
% too far apart, too close, tilted initial configurations etc.
The relative interlayer orientation was chosen based on clues from diffraction
data as well as the various known stacking modes of benzene and substituted 
benzene rings: sandwiched, parallel-displaced and T-shaped ~\cite{sinnokrot_estimates_2002}
(\Cref{fig:sandwiched,fig:pd,fig:tshaped}). The T-shaped 
configuration was ruled out based on the inconsistency of its $\approx$ 
5 \angstrom equilibrium stacking distance ~\cite{sinnokrot_estimates_2002}. The system's preference
towards the sandwiched vs. parallel displaced stacking modes was left
to be explored. Visualization of each configuration 
(\Cref{fig:sandwichedlayers,fig:offsetlayers}) suggests 
entropic differences based on the way the tails are able to pack.
In the sandwiched configuration, all tails start out directly on top 
of each other which may prevent closely stacked benzene rings. In the offset
configuration, the tails in between each other which may allow layers
to come together in a compact way. This difference may explain, in part,
which stacking mode is more favorable.

@@

An equilibration scheme with position restraints placed on aromatic rings
prevents unrealistic jumps during early equilibration steps. Restraints
fix monomer head groups in the sandwiched or parallel-displaced 
configurations while allowing monomer tails to settle. Doing so also mitigates 
system dependence on initial monomer configuration. Restrained
equilibrations are run in the NVT ensemble. Every 50 ps, force constants 
are reduced by the square root of its previous value, starting from 
1e6 KJ mol\textsuperscript{-1} nm\textsuperscript{-2}. Once the force constant
is below 10 KJ mol\textsuperscript{-1} nm\textsuperscript{-2}, the 
restraints are slowly released until there is no more restraining 
potential. The resulting unrestrained structure is allowed to 
equilibrate in the NPT ensemble for 400 - 500 ns.

@@

@@

Simulated X-ray diffraction patterns were generated based on atomic
coordinates to give a deeper understanding of the pore structure and
spacing. All atomic coordinates were simulated as gaussian spheres of
electron density  corresponding to each atom's atomic number. A three
dimensional fourier transform (FT) of the array of electron density 
results in a 3 dimensional structure factor representing the unit cell
in reciprocal space. We perform an angular average of the structure factor
about the z axis to generate a 2D cross section close to what one would see
experimentally. We matched experimental 2D WAXS patterns by iterative
improvement of our choice of initial structure and equilibration procedure. 
% Replace with a description written by Joe eventually

>>>>

Initial guesses for the remaining structural parameters were chosen
based on experimental data and treated as variables during model
development. The distance between pores was based on experimental SAXS
data for this system \cite{feng_thin_2016}. Initial configuration pore
spacing was chosen to be 10 \% larger than the experimental value to reduce
unintended repulsions resulting from a tightly packed initial configuration. 
The layer spacing was based on experimental 2D WAXS data which shows
reflections corresponding to features spaced 3.7 \angstrom apart. It has 
been hypothesized that the features are present due to pi-pi
interactions between stacked aromatic rings \cite{feng_scalable_2014}. 
Our simulations tend to equilibrate to a wider interlayer spacing of $\approx$ 4.1 \angstrom,
which inspired separate systems starting with layer spacings greater than
 4 \angstrom. We estimate the pore radius to be 0.6 nm based on past TEM 
images and size exclusion rejection data\cite{feng_scalable_2014,
feng_thin_2016,zhou_supported_2005}. Comparing a geometric
measurement of pore size taken from an atomistic model, to a less
precise, experimentally derived pore size estimate, will give ambiguous
results. When constructing pores, we chose the carboxylate carbon from
the monomer head group as a reference atom, and placed it a distance r from the
pore center, where r is the pore radius. We will not make direct comparisons
of pore radius between our model and experiment to avoid the ambiguity. 
% There is plenty I can say about initial configuration. Starting layers
% too far apart, too close, tilted initial configurations etc.
The relative interlayer orientation was chosen based on clues from diffraction
data as well as the various known stacking modes of benzene and substituted 
benzene rings: sandwiched, parallel-displaced and T-shaped ~\cite{sinnokrot_estimates_2002}
(\Cref{fig:sandwiched,fig:pd,fig:tshaped}). The T-shaped 
configuration was ruled out based on the inconsistency of its $\approx$ 
5 \angstrom equilibrium stacking distance ~\cite{sinnokrot_estimates_2002}. The system's preference
towards the sandwiched vs. parallel displaced stacking modes was left
to be explored. Visualization of each configuration 
(\Cref{fig:sandwichedlayers,fig:offsetlayers}) suggests 
entropic differences based on the way the tails are able to pack.
In the sandwiched configuration, all tails start out directly on top 
of each other which may prevent closely stacked benzene rings. In the offset
configuration, the tails are placed in between each other which may allow layers
to come together in a compact way. This difference may explain, in part,
which stacking mode is more favorable.

@@

An equilibration scheme with position restraints placed on aromatic rings
prevents unrealistic jumps during early equilibration steps. Restraints
fix monomer head groups in the sandwiched or parallel-displaced 
configurations while allowing monomer tails to settle. Doing so also mitigates 
system dependence on initial monomer configuration. Restrained
equilibrations are run in the NVT ensemble. Every 50 ps, we reduce the force constants 
by the square root of its previous value, starting from 
1e6 KJ mol\textsuperscript{-1} nm\textsuperscript{-2}. Once the force constant
is below 10 KJ mol\textsuperscript{-1} nm\textsuperscript{-2}, the 
restraints are slowly released until there is no more restraining 
potential. The resulting unrestrained structure is allowed to 
equilibrate in the NPT ensemble for 400 - 500 ns.

@@

Simulated X-ray diffraction patterns were generated based on atomic
coordinates for a direct experimental comparison. All atomic coordinates
were simulated as gaussian spheres of electron density corresponding to
each atom's atomic number. A three dimensional fourier transform 
(FT) of the array of electron density results in a three dimensional structure
factor whicih represents the unit cell in reciprocal space. We perform an
angular average of the structure factor about the z axis to generate a
2D cross section close to what one would see experimentally. We matched 
experimental 2D WAXS patterns by iterative improvement of our choice of 
initial structure and equilibration procedure. 
% Replace with a description written by Joe eventually

@@
